### 2016-10-13 Scripts/0_Pretest_SI_Classification/Pretest_Summary.Rmd
renamed script files with order of operations. begining to amend file paths in scripts.

---
title: "Pretest Summary"
author: "Lacey Etzkorn"
date: "September 23, 2016"
output: 
      html_document:
            toc: TRUE
---

```{r setup, echo=FALSE}
### Set WD
      setwd("/Users/etzkornlacey15/Documents/Reddit_Suicide/Pretest/")
      library(dplyr)
### Load Pretest Data
      pretest.lacey <- read.csv(file="lacey_classified_pretest.csv",
                                stringsAsFactors = F)
      pretest.linda <- read.csv(file="linda_classified_pretest.csv", 
                                stringsAsFactors = F) %>%
            select(X, suicidal, suicidal.past, mental.disorder)
      pretest <- merge(pretest.lacey, pretest.linda, 
                       by = c("X"), suffixes=c(".lacey", ".linda"))
### Correct Misspecified Values
      bad0 <- c("00", "o", "")
      pretest$suicidal.lacey <- as.numeric(ifelse(pretest$suicidal.lacey %in% bad0, "0", pretest$suicidal.lacey))
      pretest$suicidal.linda <- as.numeric(ifelse(pretest$suicidal.linda %in% bad0, "0", pretest$suicidal.linda))
      pretest$suicidal.past.lacey <- as.numeric(ifelse(pretest$suicidal.past.lacey %in% bad0, 
                                            "0", pretest$suicidal.past.lacey))
      pretest$suicidal.past.linda <- as.numeric(ifelse(pretest$suicidal.past.linda %in% bad0, 
                                            "0", pretest$suicidal.past.linda))
```

# Correct Mistakes

```{r}
pretest <- mutate(pretest, 
                  suicidal = suicidal.linda + suicidal.lacey > 0,
                  suicidal.past = suicidal.past.linda + suicidal.past.lacey > 0,
                  mental.disorder = mental.disorder.linda + mental.disorder.lacey > 0) 
### Current
      pretest$suicidal[38] <- T
      pretest$suicidal[46] <- T 
      pretest$suicidal[55] <- F 
      pretest$suicidal[76] <- T 
      pretest$suicidal[95] <- T 
      pretest$suicidal[90] <- T

### Past
      pretest$suicidal.past[4] <- F 
      pretest$suicidal.past[31] <- F 
      pretest$suicidal.past[60] <- T 
      pretest$suicidal.past[95] <- T 
      pretest$suicidal.past[104] <- T

### Mental Condition
      pretest$mental.disorder[8] <- T
      pretest$mental.disorder[26] <- F
      pretest$mental.disorder[30] <- T
      pretest$mental.disorder[49] <- T
      pretest$mental.disorder[76] <- F
      pretest$mental.disorder[98] <- T
      pretest$mental.disorder[102] <- T
```

# Check Blanks

```{r}
prop.table(table(pretest$text ==""))
pretest <- filter(pretest, !pretest$text=="")
```

# Tabulate by Person

```{r}
      with(pretest, table(suicidal.linda, suicidal.lacey))
      with(pretest, table(suicidal.past.linda, suicidal.past.lacey))
      with(pretest, table(mental.disorder.linda, mental.disorder.lacey))
```

### Look at Posts by Reddit

```{r}

### Tabulate by Subreddit
      with(pretest, table(subreddit, suicidal))
      with(pretest, table(subreddit, suicidal.past))
      with(pretest, table(subreddit, mental.disorder))
      with(pretest, table(subreddit))

### Tabulate Outcomes
      with(pretest, table(suicidal, mental.disorder))
### Distribution of dates
      hist(pretest$date)
```

### Save Data

```{r}
pretest <- 
pretest %>% 
select(title,subreddit,author,date,text,suicidal,suicidal.past)
save(pretest, file="Pretest_Cleaned_Data.RData")
```