### 2016-10-29 Supplement.Rmd
we did it!!!

---
title: "Supplemental Details"
author: "Lacey Etzkorn"
date: "October 26, 2016"
output: 
      html_document:
            toc: TRUE
---

```{r setup, message=FALSE, echo=FALSE}
library(ggplot2)
library(dplyr)
library(chron)
library(lubridate)
options(width = 200)
```

# Data Collection

The goal of our analysis is to estimate the expected number of drop-outs attributable to SI among June /r/Depression authors. The process of collecting the necessary data and estimating this parameter can be broken down into 4 modular steps:

1. **Collect all posts and comments from /r/depression in June 2016. **
We use the package subredditarchive, and modify the existing code from the python script here: https://github.com/peoplma/subredditarchive. The modified code that we used to scrape the June data can be found in 
[Scripts/0_Scrape_June_Data.py](Scripts/0_Scrape_June_Data.py).
We parse the json files for each thread for the main post and comments. For each, grab information on author, title (if applicable), text, creation date, name/link id, etc. 
(See code: [Scripts/1_Read_Json_Files.R](Scripts/1_Read_Json_Files.R))

2. **Combine and Classify June.**
Combine post and comment data into one data frame where the level of each observation is either a post or comment. Classify whether each post or comment from contained evidence of SI. (See [SI Classification](#si-classification)). 
Label authors as SI-positive if they had any posts or comments with SI in June 2016 /r/depression. Otherwise they are SI-negative. List unique SI-positive authors.
(See code: [Scripts/2_Combine_Classify_June_Posts_Comments.R](Scripts/2_Combine_Classify_June_Posts_Comments.R))

```{r, message=FALSE}
load(file = "Data/June_Depression_Sample/June_Posts_Classified.RData")
user.data$author[user.data$author == "\\[deleted\\]"] = NA
user.data$text[grepl("\\[deleted\\]",user.data$text)] = NA
summarise(user.data,
          total_posts = n(),
          total_si = sum(si),
          total_authors_si = length(unique(author[si])),
          unique_authors = length(unique(author)),
          unique_threads = length(unique(thread_id)))
authors = with(user.data, unique(author[si])) %>% na.omit()
authors.noSI = with(user.data, unique(author[!si])) %>% na.omit() %>% sample(500)
length(authors); length(authors.noSI)
rm(authors, authors.noSI, user.data)
```

3. **Scrape Author Data.**
Collect all available post and comment data for every SI-positive authors. Collect all available post and comment data for a similar-sized simple random sample of SI-negative authors. We collected ours at 4pm on October 10, 2016. 
We user this code to scrape the author data: [Scripts/3.1_Scrape_User_Data.R](Scripts/3.1_Scrape_User_Data.R),
[Scripts/3.2_Scrape_NoSI_User_Data.R](Scripts/3.2_Scrape_NoSI_User_Data.R). 
It takes nearly 24 hours to run. 
We use this code to classify the comments as SI-expressive:
[Scripts/4.1_Classify_SI_Author_Data.R](Scripts/4.1_Classify_SI_Author_Data.R),
[Scripts/4.2_Classify_NoSI_Author_Data.R](Scripts/4.2_Classify_NoSI_Author_Data.R).
This takes ~30 minutes to run. 
We use this code to combine the data and format it for plots, as well as summarize it for the estimation: 
[Scripts/5_Combine_Authors.R](Scripts/5_Combine_Authors.R).

```{r}
load("Data/SI_Author_Sample/June_Authors_Combined.RData")
user.data %>%
      group_by(si.author) %>%
      summarise(
          n_author = length(unique(author)),
          n_posts = n(),
          posts_per_author = n_posts / n_author,
          n_posts_si = sum(class, na.rm=T))
```

4. **Drop-Out Estimation.** Estimate the expected number of drop-outs attributable to SI. We save the posterior samples to plot in the paper.
(See code: [Scripts/6_MC_Estimation.R](Scripts/6_MC_Estimation.R))

# SI Classification

We implement an algorithm that searches post text for at least one phrase from a reference list of SI-expressive phrases. We generated our reference list by programatically expanding a sample of SI-expressive phrases which were hand-picked from 750 random June /r/Depression comments.

The sample of SI-expressive comments from the 750 posts is here: 
[Data/Training/suicidal_phrases.txt](Data/Training/suicidal_phrases.txt)

We first organize the phrases into linguistic patterns. There seemed to be twelve types of phrases, or phrase "recipes", people used to express suicidal ideation. Each "recipe" is made up of ingredients, or smaller phrases/words that can be exchanged to convey the same meaning. The recipes are listed below. 
The ingredients are listed in the file [Scripts/99_Suicide_Syntax.txt](Scripts/99_Suicide_Syntax.txt).

Phrase Recipes:

* [I am] [intending to] [suicide verb]

* [I am] [thinking about] [suicide noun]

* I [intend to] [suicide verb]

* I [suicide verb]

* [I am] [suicide adjective]

* Why not [suicide verb]

* I [think about] [suicide noun]

* me [suicide verb]

* me [suicide noun]

* I [oppose] [live noun]

* I [am opposed to] [living]

##### SI Classification: PPV & NPV

We use the script [Scripts/99_Assign_Test_Values.R](Scripts/99_Assign_Test_Values.R) to sample 200 posts and hand-classify them as having true SI or not.

```{r ppv, message = FALSE, warning=FALSE}
load(file = "Data/Test/Test_Classified.Rdata")
colnames(test.sample)
with(test.sample, table(si, suicidal))
library(Hmisc)
binconf(c(88, 96), c(100, 100), alpha=0.05, method=c("exact"))
```

We can see that the positive predictive value is 88% and the negative predictive value is 96%.